---
title: "Reproducible research 1st proejct"
author: "XC"
date: "8/12/2017"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Introduction

t is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the ¡°quantified self¡± movement ¨C a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

## Load and process the data

```{r, eval=TRUE,warning=FALSE,echo=TRUE,message=FALSE}
library(plotly)
setwd("D:/R/reproducible research")
activitydata <- read.csv("activity.csv", header = TRUE, sep = ",", na.strings = "NA")

activitydata$date <- as.Date(activitydata$date, format = "%Y-%m-%d")

activitydata$interval <- factor(activitydata$interval)

dataWONA <- is.na(as.character(activitydata$steps))
dataWONA <- activitydata[!dataWONA,]

```

## Pareto plot and calculae mean/median

You can also embed plots, for example:

```{r eval=TRUE,warning=FALSE,echo=TRUE,message=FALSE}
stepsperday <- aggregate(steps ~ date, data = dataWONA, sum)

hist(as.numeric(stepsperday$steps), breaks = 15, col = "blue", xlab = "Number of Steps", main= "Total number of steps taken per day")


mean(stepsperday$steps)
median_WONA<-median(stepsperday$steps)
median_WONA
```

## Time series plot of the average number of steps taken

```{r eval=TRUE,warning=FALSE,echo=TRUE,message=FALSE}
steps_per_interval <- aggregate(dataWONA$steps, by=list(interval=dataWONA$interval), FUN=mean)
colnames(steps_per_interval) <- c("interval", "average_steps")
  plot_ly(x=steps_per_interval$interval,y=steps_per_interval$average_steps,mode='lines')
  
#max number of steps within 5-minutes interval
  max_steps <- max(steps_per_interval$average_steps)
  max_steps

#interval which contains the max number of steps
  
  intervale_max_steps<-steps_per_interval[which.max(steps_per_interval$average_steps),]$interval
intervale_max_steps
  
```
## Replace missing data with interval mean and calcualte mean/median

```{r eval=TRUE,warning=FALSE,echo=TRUE,message=FALSE}

NA_index <- which(is.na(as.character(activitydata$steps)))
complete_data <- activitydata

complete_data[NA_index, ]$steps	<-unlist(lapply(NA_index, FUN=function(NA_index){
                steps_per_interval[activitydata[NA_index,]$interval==steps_per_interval$interval,]$average_steps
                }))
steps_each_day_complete <- aggregate(steps ~ date, data = complete_data, sum)
colnames(steps_each_day_complete) <- c("date", "steps")

hist(as.numeric(steps_each_day_complete$steps), breaks = 20, col = "red", xlab = "Number of Steps", main= "Histogram of the total number of steps taken each day")

# number of stes per day (mean and median)

mean(steps_each_day_complete$steps)
median_repl<-median(steps_each_day_complete$steps)
median_repl
```
two data sets (with or without replacement) have same mean and slightly different median(`r median_WONA`, `r median_repl`) 

## Plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r eval=TRUE,warning=FALSE,echo=TRUE,message=FALSE}

complete_data$day <- ifelse(as.POSIXlt(as.Date(complete_data$date))$wday%%6 == 
                                    0, "weekend", "weekday")
complete_data$day <- factor(complete_data$day, levels = c("weekday", "weekend"))

steps.interval= aggregate(steps ~ interval + day, complete_data, mean)

plot_ly(steps.interval, x = ~interval, y = ~steps) %>%
add_lines(color = ~day,alpha = 0.2) %>%
  add_lines(linetype=~day)
```


